A 3d curvilinear skeletonization algorithm with application to path tracing

### 2012-09-13 skeleton_and_pathtracing.tex
Petites corrections suite à la relecture de Michel
Nouvelles corrections

\def\mydist1{D_1}
\def\mydd1{\Omega_1}
\def\mydecent{\mathcal{DC}_1}
\def\dist1{d_1}
\def\Dist1#1{\mydist1(#1)}
\def\DD1#1{\mydd1(#1)}
\def\decent#1{\mydecent(#1)}
\def\cplDist1#1{\mydist1^{cc}(#1)}
\def\cplDD1#1{\mydd1^{cc}(#1)}
\def\cpldecent#1{\mydecent^{cc}(#1)}

@@

\title{A 3d curvilinear skeletonization algorithm with application to path tracing}
\author{John Chaussard\inst{1} \and Laurent No\"{e}l\inst{2} \and Venceslas Biri\inst{2} \and Michel Couprie\inst{2}}
\institute{Universit\'{e} Paris 13, Sorbonne Paris Cit\'{e}, LAGA, CNRS(UMR 7539), \\ F-93430, Villetaneuse, France \\ \email{chaussard@math.univ-paris13.fr}
\and
Universit\'{e} Paris Est, LABINFO-IGM, A3SI-ESIEE \\ 2, boulevard Blaise Pascal, Cit\'{e} DESCARTES \\ BP 99  93162 Noisy le Grand CEDEX, France \\ \email{laurent.noel@esiee.fr, v.biri@esiee.fr, michel.couprie@esiee.fr}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                        %
%                   INTRODUCTION                         %
%                                                        %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec::intro}
Path tracing algorithms are able to render photorealistic images from a scene description by simulating light propagation. The photorealistic aspect is achieved by simulating diffusion of light: elements of the scene which are illuminated by a light source can cast this light to other elements which are not directly lighted by any light source, thus giving a natural effect of light diffusion. To do so, a path tracing algorithm sends rays (of light) in the scene

The algorithm sends rays of light simulates global illumination effect: parts of the scene which are not directly illuminated by a light source the idea behind the algorithm is very elegant: for each pixel of the image, it throw rays through the pixel and let it bounce over in the scene (like a real ray of light would do). At each bouncethat will bounce on the scene to accumulate lighting exchanges. In order to get fast convergence the rays must be efficiently distributed through the scene, toward the lights. Our idea is to use a curvilinear skeleton of the void of the scene to guide the rays in the correct direction. We start by defining some useful concepts related to path tracing and we present our skeleton based path tracing algorithm.

Noise in photorealistic computer generated images. Need to accelerate convergence. Add knowledge of the geometry and topology of the scene thanks to discrete geometry. Use the skeleton. Need of a robust curvilinear skeleton that represent topology and approximate geometry. Use the cubical complex framework.

@@

We denote by $\allfaces{n}$ the set composed of all faces in $\Z^n$. Given $m \in \dinterval{0}{n}$, we denote by $\subfaces{n}{m}$ the set composed of all $m$-faces in $\Z^n$.
Let $f \in \allfaces{n}$. We set $\cell{f} = \{g \in \allfaces{n}\vert g \subseteq f\}$, and $\cellstrict{f} = \cell{f} \setminus \{f\}$. Any element of $\cell{f}$ (resp. $\cellstrict{f}$) is \myem{a face of $f$} (resp. \myem{a proper face of $f$}.
The \myem{closure} of a set of faces $X$ is the set $\closure{X}=\union\{\cell{f} \vert f \in X\}$. The set $\Compl{X}$ is $\allfaces{n} \setminus X$.

@@

The collapse operation is the basic operation for performing homotopic thinning of a complex, and consists of removing two distinct faces $(f,g)$ from a complex $X$ under the condition that they are free:

@@

\section{A parallel directional thinning based on cubical complex}

@@

Intuitively, we want our thinning algorithm to remove free faces of a complex \quotes{layer by layer} and avoid having unequal thinning of the input complex. Therefore, we want each execution of the algorithm to remove free faces located on the border of the input complex. 
We define $\border{X}$ as the set all faces belonging to a free pair for $X$. We now introduce Alg.~\ref{algo::pardircollapse_v0}, a directional parallel thinning algorithm.

@@

\section{Aspect preservation during thinning: a parameter-free method}
\label{sec::aspect_no_param}
As previously said, our goal is to use the skeleton of a scene in order to guide light rays from the camera of the scene towards the light. In order to achieve this objective, we need a skeleton which \quotes{looks like} the original scene it was calculated from.

For example, if the input is a corridor (the void of the corridor), then the skeleton should be a line following the main direction of the corridor. 
However, Alg.~\ref{algo::pardircollapse_v0} does not necessarily preserve "geometrical features" of the input object in the resulting skeleton, which end up being a single vertex.
Moreover, when one wants to preserve "geometrical information" from the original shape in the skeleton, then it is important to not retain "too much information" during thinning in order to avoid noisy branches in the skeleton.

@@

\subsection{The lifespan of a face}
In the following, we define additional functions in the cubical complex, related to the thinning process (Alg.~\ref{algo::pardircollapse_v0}), which are essential for the filtering step of the skeleton. The first one we present is the \myem{death date} of a face.

@@

\begin{definition}
Let $f \in \complex{n}{X}$, the \myem{birth date of $f$ in $X$}, denoted by $\birth{f}{X}$, is the smallest integer $b$ such that either $f$ is a facet of $ParDirCollapse(X, \emptyset, b)$, or either $f \notin ParDirCollapse(X, \emptyset, b)$.
\end{definition}

The birth date indicates how many layers of free faces must be removed from $X$ with Alg.\ref{algo::pardircollapse_v0} before transforming $f$ into a facet of $X$ (we consider a face "lives" when it is a facet).
Finally, we can define the \myem{lifespan} of a face :

@@

The lifespan of a face $f$ of $X$ indicates how many "rounds" this face "survives" as a facet in $X$, when removing free faces with Alg.~\ref{algo::pardircollapse_v0}, and is a good indicator of how important a face can be in an object. Typically, the higher the lifespan is, and the more representative of an object's visual feature the face is. The lifespan, sometimes called \myem{saliency}, was used in \cite{Liu2009} (with the name "medial persistence") in order to propose a thinning algorithm in cubical complexes based on two parameters.

@@

Two points $x,y \in \Zset^n$ are \myem{$1$-neighbours} if the Euclidean distance between $x$ and $y$ is equal or inferior to $1$ (also called direct neighbours). A \myem{$1$-path from $x$ to $y$} is a sequence $\mathcal{C}=(z_0,...,z_k)$ of points of $\Zset^n$ such that $z_0=x$, $z_k=y$, and for all $j \in [1;k]$, $z_j$ and $z_{j-1}$ are $1$-neighbours. The length of $\mathcal{C}$ is $k$. Remark that $d_1$ is indeed the $6$-distance in the 3d voxel framework.

We set $\dist1(x,y)$ as the length of the shortest $1$-path from $x$ to $y$. Let $S \subset \Zset^n$, we set, for all $x \in \Zset^n$, the map $\Dist1{S}: \Zset^n \rightarrow \Nset$ such that $\Dist1{S}(x) = \displaystyle \min_{y \in \Compl{S}} \dist1(x,y)$.

@@

\subsection{Parameter-free thinning based on the lifespan, opening function and decenterness}
As previously said, we add edges to the set $W$ of Alg.~\ref{algo::pardircollapse_v0} in order to retain, in the resulting curvilinear skeleton, important edges from the original object. Given a cubical complex $X$, if an edge of $X$ has a high decenterness value for $X$, then it is probably located too close to the border of $X$ and does not represent an interesting visual feature to preserve.
%TODO ref vers image de decenterness
On the other hand, if an edge has a high lifespan for $X$, then it means it was not removed quickly, after becoming a facet, by the thinning algorithm and might represent some precious visual information on the original object. An idea would be to keep, during thinning, all edges whose lifespan is superior to the decenterness value. Unfortunately, this strategy produces skeletons with many spurious branches in surfacic areas of the original object.

>>>>

\def\mydist1{D}
\def\mydd1{\Omega}
\def\mydecent{\mathcal{DC}}
\def\dist1{d}
\def\Dist1#1{\mydist1(#1)}
\def\DD1#1{\mydd1(#1)}
\def\decent#1{\mydecent(#1)}
\def\cplDist1#1{\mydist1^{cc}(#1)}
\def\cplDD1#1{\mydd1^{cc}(#1)}
\def\cpldecent#1{\mydecent^{cc}(#1)}

@@

\title{A 3d curvilinear skeletonization algorithm with application to path tracing}
\author{John Chaussard\inst{1} \and Laurent No\"{e}l\inst{2} \and Venceslas Biri\inst{2} \and Michel Couprie\inst{2}\thanks{This work has been partially supported by the \quotes{ANR-2010-BLAN-0205 KIDICO} project}}
\institute{Universit\'{e} Paris 13, Sorbonne Paris Cit\'{e}, LAGA, CNRS(UMR 7539), \\ F-93430, Villetaneuse, France \\ \email{chaussard@math.univ-paris13.fr}
\and
Universit\'{e} Paris Est, LIGM, A3SI-ESIEE \\ 2, boulevard Blaise Pascal, Cit\'{e} DESCARTES \\ BP 99  93162 Noisy le Grand CEDEX, France \\ \email{laurent.noel@esiee.fr, v.biri@esiee.fr, michel.couprie@esiee.fr}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                        %
%                   INTRODUCTION                         %
%                                                        %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec::intro}
Path tracing algorithms are able to render photorealistic images from a scene description by simulating light propagation. The photorealistic aspect is achieved by simulating diffusion of light: elements of the scene which are illuminated by a light source can cast part of the light they receive. Consequently, parts of the scene which are not directly illuminated by any light source actually receive light, thus giving a realistic effect of light diffusion to the resulting image.

A path tracing algorithm sends rays (of light) from the camera (the viewpoint of the scene), and let them bounce around the scene until they encounter a light source. When a ray hit a point of an object (like a wall), it gathers some information on the object's material. The ray is then reemitted in a random direction. Once a ray reaches a light source, the information it gathered through its journey in the scene is averaged in order to decide of the colour of the first impact point. This color is then allocated to one of the pixel of the image, and the process is repeated for each pixel of the output image. To avoid \quotes{unrealistic} results due to an unusual reemission direction of a ray, the algorithm actually sends, for each pixel of the image, many rays: the more rays, the longer the algorithm will take to produce an image, but the more realistic the result will look. A path tracing can be seen as a random walk process through the scene.

The bouncing direction of a ray depends on a function associated to the texture of the hit point: this probability distribution function gives, for a given incoming direction, the probability of reemitting the ray in another directions. The bouncing direction is randomly chosen among the possible values of this function. Choosing only one bouncing direction produces a noise in the resulting image (due, as previously said, to points where unusual directions were chosen): this noise is naturally reduced by shooting many rays for each image pixel (however, it will remain).

In order to avoid infinite bouncing of rays in the scene, a limit is set for the number of bounces a ray can do. If a ray does not reach any light source after a given number of bounces, it is eliminated from the scene. A lost ray is as a waste of computation time, and minimizing this loss would speed up the algorithm.

Our proposition is to reduce both the time taken by a path tracing algorithms and the noise in the output images by guiding rays towards the scene's light sources, thus reducing the number of rays \quotes{lost} during the computation. To do this, we could perform some precomputation on the void space of the scene (the medium where the rays of light propagate) in order to have, later in the algorithm, a clue on towards which direction a ray should bounce in order to meet a light source. Such precomputation can take time and become more expensive than the original algorithm.

To keep the precomputation fast, we would perform it only on a subset of points of the voids of the scene. The precomputation we want to perform relies on shortest paths and visibility tests, we therefore want to have a reduced connected set of points (topology preservation). Moreover, in order to obtain correct results, this set of points should contain geometrical information from the original scene (for example, it should possess branches propagating in the elongated parts of the scene).

A filtered curvilinear skeleton of the voids of the scene would meet all the required properties for our subset of points: it is a thin subset of points of the scene, with the same topology and retaining the main geometrical information of the scene (if the filtering step is efficient). In the following, we first present a new curvilinear skeletonization algorithm (Sec.~\ref{sec::skeletonization}) in the cubical complex framework (Sec.~\ref{sec::binary_to_cubic}), producing filtered skeletons without needing any user input (Sec.~\ref{sec::aspect_no_param}). Then, we show how to use such skeletons to enhance the results of a path tracing algorithm (see Sec.~\ref{sec::appli_pathtrace}).

@@

We denote by $\allfaces{n}$ the set composed of all faces in $\Z^n$. Given $m \in \dinterval{0}{n}$, we denote by $\subfaces{n}{m}$ the set composed of all $m$-faces in $\Z^n$.
Let $f \in \allfaces{n}$. We set $\cell{f} = \{g \in \allfaces{n}\vert g \subseteq f\}$, and $\cellstrict{f} = \cell{f} \setminus \{f\}$. Any element of $\cell{f}$ (resp. $\cellstrict{f}$) is \myem{a face of $f$} (resp. \myem{a proper face of $f$}.
The \myem{closure} of a set of faces $X$ is the set $\closure{X}=\union\{\cell{f} \vert f \in X\}$.

@@

The collapse operation is the basic operation for performing homotopic thinning of a complex, and consists of removing two distinct faces $(f,g)$ from a complex $X$ under the condition that they form a free pair:

@@

\section{A parallel directional thinning based on cubical complex}
\label{sec::skeletonization}

@@

Intuitively, we want our thinning algorithm to remove free faces of a complex \quotes{layer by layer} and to avoid having unequal thinning of the input complex. Therefore, we want each execution of the algorithm to remove free faces located on the border of the input complex. 
We define $\border{X}$ as the set all faces belonging to a free pair for $X$. We now introduce Alg.~\ref{algo::pardircollapse_v0}, a directional parallel thinning algorithm.

@@

\section{Aspect preservation during thinning: a parameter-free method}
\label{sec::aspect_no_param}
As previously said, our goal is to use the skeleton of a scene in order to guide light rays from the camera of the scene towards the light. In order to achieve this objective, we need a skeleton which captures the main geometrical features of the original scene it was calculated from.

For example, if the input is a corridor (the void of the corridor), then the skeleton should be a line following the main direction of the corridor. 
However, Alg.~\ref{algo::pardircollapse_v0} does not necessarily preserve geometrical features of the input object in the resulting skeleton, which end up being a single vertex.
Moreover, when one wants to preserve geometrical information from the original shape in the skeleton, then it is important to not retain "too much information" during thinning in order to avoid noisy branches in the skeleton.

@@

\subsection{The lifespan of a face}
In the following, we define additional functions in the cubical complex, related to the thinning process (Alg.~\ref{algo::pardircollapse_v0}), which are essential for the filtering step of the skeletonization. The first one we present is the \myem{death date} of a face.

@@

\begin{definition}
Let $f \in \complex{n}{X}$, the \myem{birth date of $f$ in $X$}, denoted by $\birth{f}{X}$, is the smallest integer $b$ such that either $f$ is a facet of $ParDirCollapse(X, \emptyset, b)$, or $f \notin ParDirCollapse(X, \emptyset, b)$.
\end{definition}

The birth date indicates how many layers of free faces must be removed from $X$ with Alg.\ref{algo::pardircollapse_v0} before transforming $f$ into a facet of $X$ (we consider that a face \quotes{lives} when it is a facet).
Finally, we define the \myem{lifespan} of a face :

@@

The lifespan of a face $f$ of $X$ indicates how many "rounds" this face "survives" as a facet in $X$, when removing free pairs with Alg.~\ref{algo::pardircollapse_v0}, and is a good indicator of how important a face can be in an object. Typically, the higher the lifespan is, and the more representative of an object's geometrical feature the face is. The lifespan, sometimes called \myem{saliency}, was used in \cite{Liu2009} (with the name "medial persistence") in order to propose a thinning algorithm in cubical complexes based on two parameters.

@@

Two points $x,y \in \Zset^n$ are \myem{$1$-neighbours} if the Euclidean distance between $x$ and $y$ is equal or inferior to $1$ (also called direct neighbours). A \myem{$1$-path from $x$ to $y$} is a sequence $\mathcal{C}=(z_0,...,z_k)$ of points of $\Zset^n$ such that $z_0=x$, $z_k=y$, and for all $j \in [1;k]$, $z_j$ and $z_{j-1}$ are $1$-neighbours. The length of $\mathcal{C}$ is $k$.

We set $\dist1(x,y)$ as the length of a shortest $1$-path from $x$ to $y$; remark that $\dist1$ is indeed the $6$-distance in the 3d voxel framework. Let $S \subset \Zset^n$, we set $\Compl{S} = \Zset^n \setminus S$, and for all $x \in \Zset^n$, the map $\Dist1{S}: \Zset^n \rightarrow \Nset$ is such that $\Dist1{S}(x) = \displaystyle \min_{y \in \Compl{S}} \dist1(x,y)$.

@@

\subsection{Parameter-free thinning based on the lifespan, opening function and decenterness}
As previously said, we add edges to the set $W$ of Alg.~\ref{algo::pardircollapse_v0} in order to retain, in the resulting curvilinear skeleton, important edges from the original object. Given a cubical complex $X$, if an edge of $X$ has a high decenterness value for $X$, then it is probably located too close to the border of $X$ and does not represent an interesting geometrical feature to preserve.
%TODO ref vers image de decenterness
On the other hand, if an edge has a high lifespan for $X$, then it means it was not removed quickly, after becoming a facet, by the thinning algorithm and might represent some precious geometrical information on the original object. An idea would be to keep, during thinning, all edges whose lifespan is superior to the decenterness value. Unfortunately, this strategy produces skeletons with many spurious branches in surfacic areas of the original object.
